---
title: "Ottawa Tax Roll"
author: "Patrick Bickerton"
date: "2025-03-30"
output: 
  prettydoc::html_pretty:
    theme: hpstr
    highlight: github
    
#output: distill::distill_article
---


## Libraries
The following libraries are required in the analysis

```{r}
# import the libraries
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(readr)
library(dplyr)
library(scales) # for currency formatting
library(glue)
library(forcats)
library(ggnewscale)    # multiple fill scales
library(moments)  #for skewness
#library(psych) #for principle component analysis with VARIMAX rotation
#library(googlesheets4) #to import custom dataset
```

#Import the Data

```{r}
# Import Revenue by Parcel Data
OttawaTaxRoll <- read_csv("QGIS/Parcel/City/Revenue by Parcel2.csv", 
                          na = c("NULL", "N/A", "", "NA"),
                          show_col_types = FALSE)

# Add Zoning Parent Group
OttawaTaxRoll <- OttawaTaxRoll %>% 
  mutate(parent_zone = case_when(
    ZONE_MAIN %in% c("R1", "R2", "R3", "R4", "R5","RM") ~ "Residential",
    ZONE_MAIN %in% c("I1","I2") ~ "Institutional",
    ZONE_MAIN %in% c("L1","L2","L3","O1") ~ "Open Space & Leisure",
    ZONE_MAIN %in% c("EP") ~ "Environmental",
    ZONE_MAIN %in% c("AM","GM","LC","MC","MD","TD","TM") ~ "Mixed Use & Commercial",
    ZONE_MAIN %in% c("IG", "IH","IL","IP") ~ "Industrial",
    ZONE_MAIN %in% c("T1","T2") ~ "Transportation",
    ZONE_MAIN %in% c("AG","ME","MR","RC","RG","RH","RI","RR","RU","VM","V1","V2","V3") ~ "Rural",
    TRUE ~ "Other"
    )
  )

# Transform Skewed Data
OttawaTaxRoll <- OttawaTaxRoll %>% 
  mutate(
    `2024 - Municipal taxes_sqrt` = sqrt(`2024 - Municipal taxes`),
    `2024 - Municipal taxes_log` = log(`2024 - Municipal taxes`),
    `2024 - Municipal taxes_log10` = log10(`2024 - Municipal taxes`),
    `2024 - Municipal taxes_inv` = 1/`2024 - Municipal taxes`
  )


# Import Ward Tax Data from City
Taxes_by_Ward_2022_2024 <- read_csv("City of Ottawa Open Data/Taxes_by_Ward_2022-2024.csv", 
                          na = c("NULL", "N/A", "", "NA"),
                          show_col_types = FALSE)

Taxes_by_Ward_2022_2024 <- Taxes_by_Ward_2022_2024 %>%
  mutate(across(
    `F2024_Current_Value_Assessment_`:`F2022_Percent_of_Total`,
    parse_number
  ))
```
## Add Normalized Data

```{r}
OttawaTaxRoll <- OttawaTaxRoll %>%
  mutate(`2024 Municipal Revenue by Hectare` = 
           `2024 - Municipal taxes` / (AREA_EPSG26918/10000))



```



# Summarize the data
M
```{r}
WardTaxSummary <- OttawaTaxRoll %>%
  group_by(WARD) %>%
  summarise_at(
    c("2024 - Total assessed value", "2024 - Municipal taxes"),
    sum,
    na.rm = TRUE
  ) %>%
    left_join(
    Taxes_by_Ward_2022_2024 %>%
      select(WARD,WARD_NAME),
    by = "WARD"
  )
WardTaxSummary <- WardTaxSummary[, c(1,4,2,3)] 


OttawaTaxRoll %>%
  filter(!is.na(`2024 Municipal Revenue by Hectare`)) %>%
  ggplot(aes(x = `2024 Municipal Revenue by Hectare`)) +
  geom_histogram(binwidth = 10000, fill = "steelblue", color = "white") +
  labs(
    title = "Histogram of 2024 Municipal Revenue by Hectare",
    x = "Tax Amount per Hectare ($)",
    y = "Count of Properties"
  ) +
  theme_minimal()

OttawaTaxRollValid <- OttawaTaxRoll %>% filter(!is.na(`2024 Municipal Revenue by Hectare`))

taxHistogram2024 <- OttawaTaxRollValid %>%
  ggplot(aes(x = `2024 Municipal Revenue by Hectare`)) +
  geom_histogram(binwidth = 10000, fill = "steelblue", color = "white") +
  labs(
    title = "Histogram of 2024 Municipal Revenue by Hectare",
    x = "Tax Amount per Hectare ($)",
    y = "Count of Properties"
  ) +
  theme_minimal()

# 
boxplot(OttawaTaxRollValid$`2024 Municipal Revenue by Hectare` ~ OttawaTaxRollValid$WARD)
# 
ggplot(OttawaTaxRollValid, aes(x = as.factor(OttawaTaxRollValid$WARD), y = `2024 Municipal Revenue by Hectare`)) +
  geom_boxplot(fill = "steelblue", outlier.color = "red", outlier.shape = 1) +
  labs(
    title = "2024 Municipal Revenue by Hectare",
    x = "Ward",
    y = "Municipal Tax ($)"
  ) +
  theme_minimal()

ggplot(OttawaTaxRoll_singlezone, aes(x = as.factor(ZONE_MAIN), y = `2024 Municipal Revenue by Hectare`)) +
  geom_boxplot(fill = "steelblue", outlier.color = "red", outlier.shape = 1) +
  labs(
    title = "2024 Municipal Revenue by Hectare ($/ha)",
    x = "Zoning Code",
    y = "Municipal Tax by Hectare ($/ha)"
  ) +
  coord_cartesian(ylim = c(0, 15000000)) +  # Adjust this upper limit as needed
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
ggplot(OttawaTaxRollValid, aes(x = AREA_EPSG26918, y = `2024 - Municipal taxes`, color = as.factor(WARD))) +
  geom_point(alpha = 0.2) +
  coord_cartesian(xlim = c(0, 25000), ylim = c(0, 150000)) +
  labs(
    title = "Parcel Area vs. 2024 Municipal Taxes by Ward",
    x = "Parcel Area (m²)",
    y = "Municipal Tax ($)",
    color = "Ward"
  ) +
  theme_minimal()

zoning_means <- OttawaTaxRoll_singlezone %>%
  group_by(ZONE_MAIN) %>%
  summarise(
    mean_municipal_tax_hectare = mean(`2024 Municipal Revenue by Hectare`, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(mean_municipal_tax_hectare))


```
# Data Coverage
## Tax by Ward
* Table. Row = Ward 
** Cols = 2024 revenue (city), revenue (Pat), dif
```{r}

DataComparison <- WardTaxSummary %>%
  left_join(
    Taxes_by_Ward_2022_2024 %>%
      select(WARD, `F2024_Current_Value_Assessment_`, `F2024_Municipal_Taxes`),
    by = "WARD"
  ) 



#Coverage of assessed value
DataComparison <- DataComparison %>%
  mutate(`coverage_AV_pct` = 
           round(`2024 - Total assessed value` / `F2024_Current_Value_Assessment_`,5))

#Coverage of municipal revenue
DataComparison <- DataComparison %>%
  mutate(`coverage_MT_pct` = 
           round(`2024 - Municipal taxes` / `F2024_Municipal_Taxes`,5))

DataComparison <- DataComparison[, c(1,4,2,3,5,6,7,8)] 

kable(DataComparison, caption = "Comparison of Municipal Taxes by Ward (2024)")

df_plot_MT_pct <- DataComparison %>%

  mutate(missing_MT_pct = 1-coverage_MT_pct) %>%
  
  # 3) turn WARD_NAME into a factor ordered by descending coverage
  mutate(
    WARD_NAME = fct_reorder(as.factor(WARD_NAME), coverage_MT_pct, .desc = TRUE)
  ) %>%
  
  # 4) now pivot to long form for stacking
  pivot_longer(
    cols      = c(coverage_MT_pct,missing_MT_pct),
    names_to  = "type",
    values_to = "percent"
  ) %>%
  
  # 5) ensure coverage is drawn first (bottom) and missing on top
  mutate(
    type = factor(type, levels = c("coverage_MT_pct", "missing_MT_pct"))
  )


ggplot(df_plot_MT_pct, aes(x = WARD_NAME, y = percent, fill = type)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(
    name = "Data Type",
    values = c("coverage_MT_pct" = "steelblue","missing_MT_pct" = "grey80"),
    labels = c("Scraped Data", "Official City Data")
  ) +
  scale_y_continuous(labels = scales::label_percent(scale = 100)) +
  labs(
    title = "Scraper Accuracy Rate",
    subtitle = "Minicipal taxes scraped as a percent of official ward totals",
    x = "Ward (Decreasing Accuracy)",
    y = "Minicipal Taxes (%)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )


df_plot_MT_abs <- DataComparison %>%
  mutate(
    missing_MT_abs = F2024_Municipal_Taxes - `2024 - Municipal taxes`,
    coverage_MT_pct = `2024 - Municipal taxes` / F2024_Municipal_Taxes,
    WARD_NAME = fct_reorder(WARD_NAME, F2024_Municipal_Taxes, .desc = TRUE)
  ) %>%
  select(WARD_NAME, `2024 - Municipal taxes`, missing_MT_abs) %>%
  pivot_longer(
    cols = c(`2024 - Municipal taxes`, missing_MT_abs),
    names_to = "type",
    values_to = "amount"
  ) %>%
  mutate(
    type = factor(type, levels = c("missing_MT_abs","2024 - Municipal taxes"))
  )

ggplot(df_plot_MT_abs, aes(x = WARD_NAME, y = amount, fill = type)) +
  geom_col(position = position_stack(reverse = FALSE)) +
  scale_fill_manual(
    name = "Data Type",
    values = c("2024 - Municipal taxes" = "steelblue", "missing_MT_abs" = "grey80"),
    labels = c("Ward Totals (Official)","Scraped")
  ) +
  scale_y_continuous(labels = label_dollar()) +
  labs(
    title = "Scraper Coverage: 2024 Municipal Taxes",
    x = "Ward (Descending Ward Total Revenue)",
    y = "Municipal Tax Revenue ($)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )



ggsave("Scraper Accuracy by Ward.png", plot = last_plot(), width = 9.5, height = 5, dpi = 200,bg="white")
```


## Parcel Coverage
```{r}
Ottawa_Parcel <- read_csv("QGIS/Parcel/City/Ottawa Parcel.csv", 
                          na = c("NULL", "N/A", "", "NA"),
                          show_col_types = FALSE)

Ottawa_Parcel <- Ottawa_Parcel %>% 
  mutate(parent_zone = case_when(
    ZONE_MAIN_major %in% c("R1", "R2", "R3", "R4", "R5","RM") ~ "Residential",
    ZONE_MAIN_major %in% c("I1","I2") ~ "Institutional",
    ZONE_MAIN_major %in% c("L1","L2","L3","O1") ~ "Open Space & Leisure",
    ZONE_MAIN_major %in% c("EP") ~ "Environmental",
    ZONE_MAIN_major %in% c("AM","GM","LC","MC","MD","TD","TM") ~ "Mixed Use & Commercial",
    ZONE_MAIN_major %in% c("IG", "IH","IL","IP") ~ "Industrial",
    ZONE_MAIN_major %in% c("T1","T2") ~ "Transportation",
    ZONE_MAIN_major %in% c("AG","ME","MR","RC","RG","RH","RI","RR","RU","VM","V1","V2","V3") ~ "Rural",
    TRUE ~ "Other"
    )
  )

OttawaParcelSummary <- Ottawa_Parcel %>%
  group_by(WARD) %>%
    summarise(
      Parcels = sum(!is.na(`ASSESSMENT_ROLL_NUMBER`)
    )
  )%>%
    left_join(
    Taxes_by_Ward_2022_2024 %>%
      select(WARD,WARD_NAME),
    by = "WARD"
  )

OttawaZoneSummary <- Ottawa_Parcel %>%
  group_by(parent_zone) %>%
    summarise(
      Parcels = sum(!is.na(`ASSESSMENT_ROLL_NUMBER`)
    )
  )

ZoneSummary <- OttawaTaxRoll %>%
  group_by(parent_zone) %>%
  summarise(
    Scraped_Parcels = sum(!is.na(`EXTRACTED_ADDRESS_SUMMARY`))
  )  %>% 
    left_join(
    OttawaZoneSummary %>%
      select(parent_zone,Parcels),
    by = 'parent_zone'
  ) %>% 
   mutate(coverage = 
           Scraped_Parcels/Parcels)

ZoneSummary <- ZoneSummary %>%
  arrange(-coverage)


df_plot_zone_pct <- ZoneSummary %>%
  mutate(
    missing = 1 - coverage,
    parent_zone = fct_reorder(as.factor(parent_zone), coverage, .desc = TRUE)   # ensure WARD is factor
  ) %>%
  pivot_longer(c(coverage, missing),
               names_to  = "type",
               values_to = "percent") %>%
  mutate(type = factor(type, levels = c("missing","coverage")))



ggplot(df_plot_zone_pct, aes(x = parent_zone, y = percent, fill = type)) +
  geom_col(position = position_stack(reverse = FALSE)) +
  scale_fill_manual(
    name = "Data Type",
    values = c("coverage" = "steelblue","missing" = "grey80"),
    labels = c("Total Parcels","Scraped Parcels")
  ) +
  scale_y_continuous(labels = scales::label_percent(scale = 100)) +
  labs(
    title = "Scraper Coverage by Zoning Group",
    x = "Zoning Group (Descending Coverage)",
    y = "Count (%)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )

ggsave("Scraper Coverage by Zoning Group.png", plot = last_plot(), width = 9.5, height = 5, dpi = 200,bg="white")

df_plot_zone <- ZoneSummary %>%
  mutate(
    missing_parcels = Parcels - Scraped_Parcels,
    parent_zone = fct_reorder(parent_zone, Parcels, .desc = TRUE)
  ) %>%
  pivot_longer(
    cols = c(Scraped_Parcels, missing_parcels),
    names_to = "type",
    values_to = "count"
  ) %>%
  mutate(
    type = factor(type, levels = c("Scraped_Parcels", "missing_parcels"))
  )



ggplot(df_plot_zone, aes(x = parent_zone, y = count, fill = type)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(
    name = "Parcel Status",
    values = c("Scraped_Parcels" = "steelblue", "missing_parcels" = "grey80"),
    labels = c("Scraped", "Uncovered")
  ) +
  scale_y_continuous(labels = comma) +
  labs(
    title = "Scraped vs. Uncovered Parcels by Zoning Code",
    x = "Zoning Code (Descending Total Parcels)",
    y = "Number of Parcels"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )

ggsave("Scraper Coverage (absolute) by Zoning Group.png", plot = last_plot(), width = 9.5, height = 5, dpi = 200,bg="white")

```

## Revenu by Area

```{r}
ggplot(OttawaTaxRoll, aes(x = `2024 - Municipal taxes_log`)) +
  geom_histogram(bins = 100, fill = "steelblue", color = "white") +
  labs(title = "Histogram of Square Root Transformed Municipal Taxes (2024)",
       x = "Log(Municipal Taxes)",
       y = "Frequency") +
  theme_minimal()

OttawaTaxRoll <- OttawaTaxRoll %>%
  mutate(
    log_revenue_per_hectare = log(`2024 Mun Revenue by Hectare`)
  )

OttawaTaxRoll_R_MC <- OttawaTaxRoll %>%
  filter(
    parent_zone == c("Residential","Mixed Use & Commercial")
  )

ggplot(OttawaTaxRoll_R_MC, aes(x = AREA_EPSG26918, y = `2024 - Municipal taxes_log10`, color = as.factor(ZONE_MAIN))) +
  geom_point(alpha = 0.2) +
  #stat_ellipse(type = "norm", level = 0.95) +
  coord_cartesian(xlim = c(0, 25000), ylim = c(0, 20)) +
  labs(
    title = "Parcel Area vs. 2024 Municipal Taxes by Ward",
    x = "Parcel Area (m²)",
    y = "Log(Municipal Tax)",
    color = "Zone"
  ) +
  theme_minimal()



OttawaTaxRoll_singlezone <- OttawaTaxRoll_R_MC %>%
  filter(!grepl(",", ZONE_MAIN)) %>%
  mutate(ZONE_MAIN = fct_reorder(ZONE_MAIN,log_revenue_per_hectare,.fun =mean, .desc= TRUE))

ggplot(OttawaTaxRoll_singlezone, aes(x = log_revenue_per_hectare)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  labs(
    title = "Distribution of Revenue per Hectare",
    x = "Revenue per Hectare",
    y = "Frequency"
  ) +
  theme_minimal()

model <- lm(log_revenue_per_hectare ~ ZONE_MAIN, data = OttawaTaxRoll_singlezone)
summary(model)

ggplot(OttawaTaxRoll_singlezone, aes(x = ZONE_MAIN, y = log_revenue_per_hectare)) +
  geom_boxplot(fill = "lightblue") +
   coord_cartesian(ylim = c(5, 16))+
    scale_x_discrete(labels = c(
    "AM" = "Arterial Mainstreet",
    "R1" = "Residential 1",
    "R2" = "Residential 2",
    "R3" = "Residential 3",
    "R4" = "Residential 4",
    "R5" = "Residential 5",
    "RM" = "Mobile Homes",
    "GM" = "General Mixed Use",
    "LC" = "Local Commercial",
    "MD" = "Mixed Use Downtown",
    "MC" = "Mixed-Use Centre",
    "TD" = "Transit Oriented Development",
    "TM" = "Traditional Mainstreet"
  )) +
  labs(
    title = "Municipal Tax Revenue by Hectare by Zoning Code",
    x = "Zoning Code (Descending)",
    y = "Log(Rev per Hectare)",
    caption = "Displayed on log scale to improve readability. Larger values appear visually closer to smaller values than in reality."
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
    axis.text.y = element_text(size = 14),
    title = element_text(size = 14),
    plot.margin = margin(t = 10, r = 30, b = 10, l = 30)
    )

ggsave("Municipal Tax Revenue by Zoning Code.png", plot = last_plot(), width = 9, height = 6, dpi = 200,bg="white")


ggplot(zoning_means, aes(x = reorder(ZONE_MAIN, mean_municipal_tax_hectare), y = mean_municipal_tax_hectare)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  scale_y_log10(labels = scales::label_comma()) +
  geom_text(
    aes(label = comma(round(mean_municipal_tax_hectare))),
    hjust = 1.1,
    vjust = .5,
    size = 4,
    color = 'white'
  ) +
  scale_x_discrete(labels = c(
    "AM" = "Arterial Mainstreet",
    "R1" = "Residential 1",
    "R2" = "Residential 2",
    "R3" = "Residential 3",
    "R4" = "Residential 4",
    "R5" = "Residential 5",
    "RM" = "Mobile Homes",
    "GM" = "General Mixed Use",
    "LC" = "Local Commercial",
    "MD" = "Mixed Use Downtown",
    "MC" = "Mixed-Use Centre",
    "TD" = "Transit Oriented Development",
    "TM" = "Traditional Mainstreet"
  ))+
  labs(
    title = "Mean Revenue Density by Zoning Code (Log Scale)",
    x = "Zoning Code",
    y = "Mean Revenue Density - Log Scale ($/ha)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

ggsave("Municipal Tax Revenue (y.axis log) by Zoning Code.png", plot = last_plot(), width = 9.5, height = 5, dpi = 200,bg="white")

ggplot(zoning_means, aes(x = reorder(ZONE_MAIN, mean_municipal_tax_hectare), y = mean_municipal_tax_hectare)) +
  geom_col(fill = "steelblue") +
  geom_text(
    aes(label = comma(round(mean_municipal_tax_hectare),100)),
    hjust = -0.1,
    vjust = 0.5,
    size = 4,
    color = "steelblue"
  ) +
  coord_flip(ylim = c(0, 2000000)) +  # Increase upper Y limit (this is now the horizontal axis)
  scale_x_discrete(labels = c(
    "AM" = "Arterial Mainstreet",
    "R1" = "Residential 1",
    "R2" = "Residential 2",
    "R3" = "Residential 3",
    "R4" = "Residential 4",
    "R5" = "Residential 5",
    "RM" = "Mobile Homes",
    "GM" = "General Mixed Use",
    "LC" = "Local Commercial",
    "MD" = "Mixed Use Downtown",
    "MC" = "Mixed-Use Centre",
    "TD" = "Transit Oriented Development",
    "TM" = "Traditional Mainstreet"
  )) +
  labs(
    title = "Mean Revenue Density by Zoning Code (city wide)",
    x = "Zoning Code",
    y = "Mean Revenue Density ($/ha)"
  ) +
  theme_minimal() +
  theme(
    plot.margin = margin(r = 40),  # Extra space for labels
    axis.text.x = element_text(size = 10)
  )

ggsave("Municipal Tax Revenue by Zoning Code.png", plot = last_plot(), width = 9.5, height = 5, dpi = 200,bg="white")
```


## Zoning Coverage
```{r}
OttawaZoneSummary <- Ottawa_Parcel %>%
  group_by(ZONE_MAIN_major) %>%
    summarise(
      Count = sum(!is.na(`ASSESSMENT_ROLL_NUMBER`)
    )
  ) %>% 
    mutate(
      ZONE_MAIN = ZONE_MAIN_major
  ) 

ZoneSummary <- OttawaTaxRoll %>%
  group_by(ZONE_MAIN) %>%
  summarise(
    Count_scraped = sum(!is.na(`EXTRACTED_ADDRESS_SUMMARY`))
  )  %>% 
    left_join(
    OttawaZoneSummary %>%
      select(ZONE_MAIN,Count),
    by = "ZONE_MAIN"
  ) 

#Coverage of Parcel Zone Scrape
ZoneSummary <- ZoneSummary %>%
  mutate(`Coverage - Zone Scrape` = 
           round(Count_scraped / Count,3)*100)%>%
  filter(!is.na(`Coverage - Zone Scrape`))


kable(ZoneSummary, caption = "Comparison of Zoning Coverage")

```


## Revenues Overall city
* Table. Row = Ward
** Cols = sum (assessed val, mun taxes, education taxes, other taxes, solid waste, total taxes), sum (par area), mun_rev/ha

```{r}
WardTaxSummary <- OttawaTaxRoll %>%
  group_by(WARD) %>%
  summarise_at(
    c("2024 - Total assessed value", "2024 - Municipal taxes","2024 - Education taxes", "2024 - Other taxes" ,"2024 - Solid waste fees","2024 - Total taxes levied","AREA_EPSG26918"),
    sum,
    na.rm = TRUE
  ) %>%
    left_join(
    Taxes_by_Ward_2022_2024 %>%
      select(WARD,WARD_NAME),
    by = "WARD"
  )

WardTaxSummary <- WardTaxSummary %>%
  mutate(`2024 - Municipal Tax by Hectare` = `2024 - Municipal taxes` / (AREA_EPSG26918/10000))

WardTaxSummary <- WardTaxSummary[, c(1,9,2,3,4,5,6,7,8,10)]

kable(WardTaxSummary, caption = "Summary of Ward Taxes")

```


** staked bar on tax $ (y axis), group by category, by ward (x axis). Raw dollar amount, not normalized by parcel area.
** box plot (x = zoning), y = mun rev by ha)
```{r}
MunicipalRevByZone <- ggplot(OttawaTaxRoll, aes(x = as.factor(OttawaTaxRoll$ZONE_MAIN), y = `2024 - Municipal taxes`)) +
  geom_boxplot(fill = "steelblue", outlier.color = "red", outlier.shape = 1) +
  labs(
    title = "2024 Municipal Revenue by Hectare",
    x = "ZONE_MAIN",
    y = "Municipal Tax ($)"
  ) +
  theme_minimal()

MunicipalRevByZone

```

# Top Ten
## Assessed Value
top 10 (summarize with table, Address Summary, Parcel, [metric], ZONE_MAIN, PAR_AREA)
** assessed value
```{r}
Top10AssedVal <- OttawaTaxRoll %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, `2024 - Total assessed value`,`2024 - Municipal taxes`,`2024 Municipal Revenue by Hectare`)

kable(Top10AssedVal, caption = "Top 10 City Properties by Assessed Value")


```
## Municipal Revenue
```{r}
Top10MunRev <- OttawaTaxRoll %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, `2024 - Total assessed value`,`2024 - Municipal taxes`,`2024 Municipal Revenue by Hectare`)


kable(Top10MunRev, caption = "Top 10 City Properties by Municipal Revenue")

```

## Revenue by Hectare 
```{r}
Top10MunRevbyHa <- OttawaTaxRoll %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, `2024 - Total assessed value`,`2024 - Municipal taxes`,`2024 Municipal Revenue by Hectare`)


kable(Top10MunRevbyHa, caption = "Top 10 City Properties by Municipal Revenue by Hectare")


```

# Bottom Ten
## Assessed Value
bottom 10 (summarize with table, Address Summary, Parcel, [metric], ZONE_MAIN, PAR_AREA)
** assessed value
```{r}
Bottom10AssedVal <- OttawaTaxRoll %>%
  arrange(`2024 - Total assessed value`) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, `2024 - Total assessed value`,`2024 - Municipal taxes`,`2024 Municipal Revenue by Hectare`)

kable(Bottom10AssedVal, caption = "Bottom 10 City Properties by Assessed Value")


```
## Municipal Revenue
```{r}
Bottom10MunRev <- OttawaTaxRoll %>%
  arrange(`2024 - Municipal taxes`) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, `2024 - Total assessed value`,`2024 - Municipal taxes`,`2024 Municipal Revenue by Hectare`)


kable(Bottom10MunRev, caption = "Bottom 10 City Properties by Municipal Revenue")

```

## Revenue by Hectare 
```{r}
Bottom10MunRevbyHa <- OttawaTaxRoll %>%
  arrange(`2024 Municipal Revenue by Hectare`) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, `2024 - Total assessed value`,`2024 - Municipal taxes`,`2024 Municipal Revenue by Hectare`)


kable(Bottom10MunRevbyHa, caption = "Bottom 10 City Properties by Municipal Revenue by Hectare")
```

# Ward Analysis
## General - Assessed Value

```{r}


generate_top10_table <- function(ward_number) {
  # Filter and prepare the data
  top10 <- OttawaTaxRoll %>%
    filter(WARD == ward_number) %>%
    arrange(desc(`2024 - Total assessed value`)) %>%
    slice_head(n = 10) %>%
    select(
      EXTRACTED_ADDRESS_SUMMARY,
      ZONE_MAIN,
      `2024 - Total assessed value`,
      `2024 - Municipal taxes`,
      `2024 Municipal Revenue by Hectare`
    ) %>%
    mutate(
      `2024 - Total assessed value` = dollar(`2024 - Total assessed value`),
      `2024 - Municipal taxes` = dollar(`2024 - Municipal taxes`),
      `2024 Municipal Revenue by Hectare` = dollar(`2024 Municipal Revenue by Hectare`)
    )
  
  # Create and return the styled table
  kbl(
    top10,
    caption = glue("Top 10 Properties in Ward {ward_number} by Assessed Value"),
    align = "l"
  ) %>%
    kable_styling(full_width = FALSE)
  
}
for (ward in 1:24) {
  print(generate_top10_table(ward))
}
```

## General - Municipal Revenue

```{r}


generate_top10_table <- function(ward_number) {
  # Filter and prepare the data
  top10 <- OttawaTaxRoll %>%
    filter(WARD == ward_number) %>%
    arrange(desc(`2024 - Municipal taxes`)) %>%
    slice_head(n = 10) %>%
    select(
      EXTRACTED_ADDRESS_SUMMARY,
      ZONE_MAIN,
      `2024 - Total assessed value`,
      `2024 - Municipal taxes`,
      `2024 Municipal Revenue by Hectare`
    ) %>%
    mutate(
      `2024 - Total assessed value` = dollar(`2024 - Total assessed value`),
      `2024 - Municipal taxes` = dollar(`2024 - Municipal taxes`),
      `2024 Municipal Revenue by Hectare` = dollar(`2024 Municipal Revenue by Hectare`)
    )
  
  # Create and return the styled table
  kbl(
    top10,
    caption = glue("Top 10 Properties in Ward {ward_number} by Municipal Tax Levy"),
    align = "l"
  ) %>%
    kable_styling(full_width = FALSE)
  
}
for (ward in 1:24) {
  print(generate_top10_table(ward))
}
```

## General - Municipal Rev by Hectare Value

```{r}


generate_top10_table <- function(ward_number) {
  # Filter and prepare the data
  top10 <- OttawaTaxRoll %>%
    filter(WARD == ward_number) %>%
    arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
    slice_head(n = 10) %>%
    select(
      EXTRACTED_ADDRESS_SUMMARY,
      ZONE_MAIN,
      `2024 - Total assessed value`,
      `2024 - Municipal taxes`,
      `2024 Municipal Revenue by Hectare`
    ) %>%
    mutate(
      `2024 - Total assessed value` = dollar(`2024 - Total assessed value`),
      `2024 - Municipal taxes` = dollar(`2024 - Municipal taxes`),
      `2024 Municipal Revenue by Hectare` = dollar(`2024 Municipal Revenue by Hectare`)
    )
  
  # Create and return the styled table
  kbl(
    top10,
    caption = glue("Top 10 Properties in Ward {ward_number} by Municipal Revenue by Hectare"),
    align = "l"
  ) %>%
    kable_styling(full_width = FALSE)
  
}
for (ward in 1:24) {
  print(generate_top10_table(ward))
}
```

## 1 - Orleans East
```{r}
Ward <- 1  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 2
```{r}
Ward <- 2  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 3 
```{r}
Ward <- 3  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 4
```{r}
Ward <- 4  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 5
```{r}
Ward <- 5  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 6
```{r}
Ward <- 6  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```

## 7
```{r}
Ward <- 7  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 8
```{r}
Ward <- 8  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`,`AREA_EPSG26918`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`,`AREA_EPSG26918`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`,`AREA_EPSG26918`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 9
```{r}
Ward <- 9  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`,`AREA_EPSG26918`,
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`,`AREA_EPSG26918`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`,`AREA_EPSG26918`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

write.csv(Top10AssessedVal, file="Ward9TopAssedVal.csv",row.names = FALSE)
write.csv(Top10MunVal, file="Ward9TopMunRev.csv",row.names = FALSE)
write.csv(Top10RevbyHaVal, file="Ward9Top10RevbyHa.csv",row.names = FALSE)

```
## 10
```{r}
Ward <- 10  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 11
```{r}
Ward <- 101# (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 12
```{r}
Ward <- 12  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 13
```{r}
Ward <- 13  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 14
```{r}
Ward <- 14  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 15
```{r}
Ward <- 15  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 16
```{r}
Ward <- 16  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 17
```{r}
Ward <- 17  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 18
```{r}
Ward <- 18  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 19
```{r}
Ward <- 19  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 20
```{r}
Ward <- 20  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 21
```{r}
Ward <- 21  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 22
```{r}
Ward <- 22  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 23
```{r}
Ward <- 23  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```
## 24
```{r}
Ward <- 24  # (small tip: it's better style to use lowercase "ward" for variables)

Top10AssessedVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Total assessed value`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)

kable(Top10AssessedVal, caption = glue("Top 10 Properties in Ward {Ward} by Assessed Value"))

Top10MunVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 - Municipal taxes`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10MunVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Tax Levy"))

Top10RevbyHaVal <- OttawaTaxRoll %>%
  filter(WARD == Ward) %>%
  arrange(desc(`2024 Municipal Revenue by Hectare`)) %>%
  slice_head(n = 10) %>%
  select(EXTRACTED_ADDRESS_SUMMARY, ZONE_MAIN, 
         `2024 - Total assessed value`, `2024 - Municipal taxes`, 
         `2024 Municipal Revenue by Hectare`)
kable(Top10RevbyHaVal, caption = glue("Top 10 Properties in Ward {Ward} by Municipal Rev by Hectare"))

```


##Session Info
This above code was run in the following environment
```{r}
sessionInfo()
```
