---
title: "Ottawa Tax Roll"
author: "Patrick Bickerton"
date: "2025-03-30"
output: distill::distill_article
---


## Libraries
The following libraries are required in the analysis

```{r}
# import the libraries
#sumlibrary(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(readr)
library(dplyr)
library(scales) # for currency formatting
#library(psych) #for principle component analysis with VARIMAX rotation
#library(googlesheets4) #to import custom dataset
```

#Import the Data

```{r}
# Import Data
OttawaTaxRoll <- read_csv("QGIS/Parcel/Revenue by Parcel.csv", 
                          na = c("NULL", "N/A", "", "NA"),
                          show_col_types = FALSE)

Taxes_by_Ward_2022_2024 <- read_csv("City of Ottawa Open Data/Taxes_by_Ward_2022-2024.csv", 
                          na = c("NULL", "N/A", "", "NA"),
                          show_col_types = FALSE)

Taxes_by_Ward_2022_2024 <- Taxes_by_Ward_2022_2024 %>%
  mutate(across(
    `F2024_Current_Value_Assessment_`:`F2022_Percent_of_Total`,
    parse_number
  ))
```
## Add Normalized Data

```{r}
OttawaTaxRoll <- OttawaTaxRoll %>%
  mutate(`2024 Municipal Revenue by Hectare` = 
           `2024 - Municipal taxes` / (AREA_EPSG26918/10000))
```

## Summarize the data
M
```{r}
WardTaxSummary <- OttawaTaxRoll %>%
  group_by(WARD) %>%
  summarise_at(
    c("2024 - Total assessed value", "2024 - Municipal taxes"),
    sum,
    na.rm = TRUE
  ) 

DataComparison <- WardTaxSummary %>%
  left_join(
    Taxes_by_Ward_2022_2024 %>%
      select(WARD,WARD_NAME, `F2024_Current_Value_Assessment_`, `F2024_Municipal_Taxes`),
    by = "WARD"
  ) 

DataComparison <- DataComparison[, c(1,4,2,3,5,6)] 
kable(DataComparison, caption = "Full Ward Tax Comparison Table")

OttawaTaxRoll %>%
  filter(!is.na(`2024 Municipal Revenue by Hectare`)) %>%
  ggplot(aes(x = `2024 Municipal Revenue by Hectare`)) +
  geom_histogram(binwidth = 10000, fill = "steelblue", color = "white") +
  labs(
    title = "Histogram of 2024 Municipal Revenue by Hectare",
    x = "Tax Amount per Hectare ($)",
    y = "Count of Properties"
  ) +
  theme_minimal()

OttawaTaxRollValid <- OttawaTaxRoll %>% filter(!is.na(`2024 Municipal Revenue by Hectare`))

taxHistogram2024 <- OttawaTaxRollValid %>%
  ggplot(aes(x = `2024 Municipal Revenue by Hectare`)) +
  geom_histogram(binwidth = 1000, fill = "steelblue", color = "white") +
  labs(
    title = "Histogram of 2024 Municipal Revenue by Hectare",
    x = "Tax Amount per Hectare ($)",
    y = "Count of Properties"
  ) +
  theme_minimal()

boxplot(OttawaTaxRollValid$`2024 Municipal Revenue by Hectare` ~ OttawaTaxRollValid$WARD)

ggplot(OttawaTaxRollValid, aes(x = as.factor(OttawaTaxRollValid$WARD), y = `2024 Municipal Revenue by Hectare`)) +
  geom_boxplot(fill = "steelblue", outlier.color = "red", outlier.shape = 1) +
  labs(
    title = "2024 Municipal Revenue by Hectare",
    x = "Ward",
    y = "Municipal Tax ($)"
  ) +
  theme_minimal()

ggplot(OttawaTaxRollValid, aes(x = as.factor(WARD), y = `2024 Municipal Revenue by Hectare`)) +
  geom_boxplot(fill = "steelblue", outlier.color = "red", outlier.shape = 1) +
  labs(
    title = "2024 Municipal Revenue by Hectare ($/ha)",
    x = "Ward",
    y = "Municipal Tax by Hectare ($/ha)"
  ) +
  coord_cartesian(ylim = c(0, 50000)) +  # Adjust this upper limit as needed
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(OttawaTaxRollValid, aes(x = AREA_EPSG26918, y = `2024 - Municipal taxes`, color = as.factor(WARD))) +
  geom_point(alpha = 0.2) +
  coord_cartesian(xlim = c(0, 25000), ylim = c(0, 150000)) +
  labs(
    title = "Parcel Area vs. 2024 Municipal Taxes by Ward",
    x = "Parcel Area (m²)",
    y = "Municipal Tax ($)",
    color = "Ward"
  ) +
  theme_minimal()

```
## Data Coverage
* Table. Row = Ward 
** Cols = 2024 revenue (city), revenue (Pat), dif
```{r}

```

** number of parcels (raw) vs. number of parcels (Pat), dif
** number of parcels per ZONE_MAIN (raw) vs. number of parcels (Pat), dif

## Revenues Overall city
* Table. Row = Ward
** Cols = sum (assessed val, mun taxes, education taxes, other taxes, solid waste, total taxes), sum (par area), mun_rev/ha
** staked bar on tax $ (y axis), group by category, by ward (x axis). Raw dollar amount, not normalized by parcel area.
** box plot (x = zoning), y = mun rev by ha)

top 10 (summarize with table, Address Summary, Parcel, [metric], ZONE_MAIN, PAR_AREA)
** assessed value
** mun revenue
** mun revenue by ha

Bottom 10 (summarize with table, Address Summary, Parcel, [metric], ZONE_MAIN, PAR_AREA)
** assessed value (non-null)
** mun revenue (non-null)
** mun revenue by ha

## Ward Analysis
Table.columm ZONE_MAIN, count parcels, total_mun revenue, total_area, avg(mun_rev by area)

top 10 (summarize with table, Address Summary, Parcel, [metric], ZONE_MAIN, PAR_AREA)
** assessed value
** mun revenue
** mun revenue by ha

Bottom 10 (summarize with table, Address Summary, Parcel, [metric], ZONE_MAIN, PAR_AREA)
** assessed value (non-null)
** mun revenue (non-null)
** mun revenue by ha





##Session Info
This above code was run in the following environment
```{r}
sessionInfo()
```
